---
output: html_document
---
# library dpylr and tidyr
library(dplyr)
library(tidyr)

---
title: 'Data Wrangling Exercise 1: Basic Data Manipulation'
author: "LMS"
date: "10th January 2017"
output: html_document
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load refine_original.csv file into RStudio and review
refine_original <- read.csv("refine_original.csv")
```

## Introduction

The data set 'refine_original.csv' is a data frame consisting of 25 observations across the following 6 variables:

'company', 'Product.code...number', 'address', 'city', 'country', 'name'.
  
In this exercise the data will be cleaned up so that brand names are consistent for each company, the 'product.code' and 'product.number' are separated and a new variable 'product_category' will be added based upon the 'product.code'.  'Address', 'city' and 'country' will be combined into a new column 'full_address' and dummy numerical variables will be made for all company and product categories.  The clean data set will be saved and written to 'refine_clean.csv'.

Please note that the column 'product.code...number' is a changed name from the original 'refine.xlsx' where it is named 'product code / product number'.  This does not always convert to 'product.code...number' in all possible conversion from .xlsx to .csv file types and should be checked in the original 'refine_original.csv' file before use.

## step 1: clean up company brand names 
Clean 'company' column to have consistent brand names: philips, akzo, van houten and unilever (all lower case).

```{r Step 1, include=FALSE}
SP_philips <- c("Phillips", "phillips", "phllips", "phillps", "phillipS", "fillips", "phlips", "philips")
SP_akzo <- c("Akzo", "AKZO", "akz0", "ak zo", "akzo")
SP_van_houten <- c("Van Houten", "van Houten", "van houten")
SP_unilever <- c("unilver", "unilever", "Unilever", "unilever")

refine_original$company[refine_original$company %in% SP_philips] <- "philips"
refine_original$company[refine_original$company %in% SP_akzo] <- "akzo"
refine_original$company[refine_original$company %in% SP_van_houten] <- "van houten"
refine_original$company[refine_original$company %in% SP_unilever] <- "unilever"
```

## step 2: seperate refine_original$Product.code...number
Separate the 'product_code' and 'product_number' into separate columns.

```{r Step 2, include=FALSE}
refine_original <- refine_original %>% 
  separate(col = Product.code...number,into = c("product_code", "product_number"), sep = "-")
```

## step 3: add the column 'product_category'
Using the newly formed 'product_code' column make a new column for each record where:
        p = Smartphone
        v = TV
        x = Laptop
        q = Tablet

```{r Step 3, include=FALSE}
ProdCatFunc <- function(x) {
  ifelse(x == "p", "Smartphone",
         ifelse(x == "v", "TV",
                ifelse(x == "x", "Laptop",
                       ifelse(x == "q", "Tablet", NA))))
}

refine_original <- refine_original %>% mutate(product_category = ProdCatFunc(product_code))
```

## step 4: condense the 3 address columns 
Bring 'address', 'city' and 'country' into one new column called 'full_address'.

```{r Step 4, include=FALSE}
refine_original <- refine_original %>% unite(full_address, address:country, sep = ",", remove = FALSE)
```

#step 5: create dummy variables for company and product categories
Create numerical variables 0 or 1 to indicate the different companies and product categories for easy selection.

```{r Step 5, include=FALSE}
# 'company_'
refine_original <- refine_original %>%
  mutate(company_philips = ifelse(company == "philips", 1, 0)) %>%
  mutate(company_akzo = ifelse(company == "akzo", 1, 0)) %>%
  mutate(company_van_houten = ifelse(company == "van houten", 1, 0)) %>%
  mutate(company_unilever = ifelse(company == "unilever", 1, 0))

# 'product_'
refine_original <- refine_original %>%
  mutate(product_smartphone = ifelse(product_category == "Smartphone", 1, 0)) %>%
  mutate(product_tv = ifelse(product_category == "TV", 1, 0)) %>%
  mutate(product_laptop = ifelse(product_category == "Laptop", 1, 0)) %>%
  mutate(product_tablet = ifelse(product_category == "Tablet", 1, 0))
```

Finally read and write cleaned data set to 'refine_clean.csv'

``` {r Write Clean CSV, include=FALSE}
refine_clean <- refine_original

write.csv(refine_clean, file = "refine_clean.csv")
```

